### 2015-07-09 2015-compmorphon-morphver/poster.tex
"preliminary development"
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61225 72bbbca6-d526-0410-a7d9-f06f51895060
some positioning and stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61227 72bbbca6-d526-0410-a7d9-f06f51895060

@@

%\usepackage{natbib}
\usepackage[backend=biber,citestyle=apa,style=authoryear,maxbibnames=99,maxcitenames=2]{biblatex}
\bibliography{poster}
\renewcommand*{\nameyeardelim}{\addcomma\space}
\renewcommand*{\bibfont}{\footnotesize}

@@

	\headerbox{Morphological Transducers}{name=morphtrans,below=morphdesc}{
		\htwo{Morphological transducers}
		\begin{itemize}
			\item Efficient (in speed \& size) models of a language's morphology
			\item Take a surface form, and produce valid lexical form(s)
			\item Take a lexical form, and produce valid surface form(s)\vspace{-0.1ex}\\
		%\end{itemize}
		\scalebox{0.98}{
			алдым \hilitetwo{↔} \texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}}, \texttt{алд\tag{n}\tag{px1sg}\tag{nom}}
		}
		\end{itemize}

@@

		\headerbox{Example:\ Tuvan velar elision}{name=tuvanvelars,column=1,span=2}{
			\begin{multicols}{2}
				\htwo{Previous descriptions}

@@

			\columnbreak{}
				\htwo{Uanalysed forms before implementation}
					\begin{itemize}
						\item \hilitetwo{инээ}, \hiliteone{өгү}, \hilitetwo{өө}, \hilitetwo{соо}, \hiliteone{чаңы}
					\end{itemize}\vspace{-0.3ex}
					\begin{minipage}[t]{0.5\textwidth}
					\outputex{
						\begin{multicols}{4}
							1590 соонда\vphantom{ү}\\
							610 те\vphantom{ү}\\
							153 өөнүң\vphantom{ү}\\
							48 соондан\vphantom{ү}\\
							46 өөм\vphantom{ү}\\
							20 соон\vphantom{ү}\\
							18 өө\vphantom{ү}\\
							17 өөнге\vphantom{ү}\\
							9 соондагы\vphantom{ү}\\
							8 өөн\vphantom{ү}\columnbreak\\
							6 инээниң\vphantom{ү}\\
							6 соонче\vphantom{ү}\\
							6 соонга\vphantom{ү}\\
							3 инээм\vphantom{ү}\\
							3 өөнден\vphantom{ү}\\
							3 өөнде\vphantom{ү}\\
							3 өөмнүң\vphantom{ү}\\
							2 инээн\vphantom{ү}\\
							2 инээ\vphantom{ү}\\
							2 өөнче\vphantom{ү}\columnbreak\\
							2 өөңерге\vphantom{ү}\\
							2 өөңер\vphantom{ү}\\
							2 өөмнү\vphantom{ү}\\
							2 өөвүстен\vphantom{ү}\\
							2 өөвүске\vphantom{ү}\\
							2 өөвүс\vphantom{ү}\\
							1 инээңни\vphantom{ү}\\
							1 инээмниң\vphantom{ү}\\
							1 Инээм\vphantom{ү}\\
							1 инээвистиң\vphantom{ү}\\
							1 өөңнүң\vphantom{ү}\\
							1 өөң\vphantom{ү}\\
							1 өөвүсче\vphantom{ү}\\
							1 өөвүстүң\vphantom{ү}\\
							1 өөвүстү\vphantom{ү}\\
							1 өөвүсте
						\end{multicols}}
					\end{minipage}

@@

	\headerbox{Example: Categorisation}{name=adjectives,column=1,below=tuvanvelars}{
		\htwo{Unrestricted 0-derivation}
		\begin{itemize}
			\item Common overgeneralisation of adjective morphology:
			\begin{itemize}
				\item all adjectives may act like nouns (i.e.\ take noun morphology)
				\item all adjectives may be used as is as adverbs
			\end{itemize}
			\item Commonly reported in Turkic language grammars
		\end{itemize}
		%Somfai Kara (2002, pp. 28-29):\ \includegraphics[width=0.53\textwidth]{img/kara28}\\
		\textcite[28-29]{somfaikara02}:\ \includegraphics[width=0.53\textwidth]{img/kara28}\\
		\-\hfill{}\includegraphics[width=0.8\textwidth]{img/kara29}

		\htwo{Attested system}
			\begin{itemize}
				\item Some adjectives may not be substantivised
				\item Some adjectives may not be used adverbially
				\item Some adjectives do not have comparative forms
				\item …a range of “adjective classes” in most Turkic languages
			\end{itemize}
		\htwo{Implementation: proper categorisation}
			\begin{itemize}
				\item only correct forms are analysed and generated
			\end{itemize}

			\scalebox{0.83}{
				\begin{tabular}{lllll}
					\toprule
					\textbf{Type} & \textbf{Gloss} & {\small \texttt{\tag{adj}(\tag{comp})}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{subst}}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{advl}}} \\
					\midrule
					\textbf{A1} & \eng{good} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} \\
					\textbf{A2} & \eng{old} & {\qipa иске (искерәк)} & {\qipa иске (искерәк)} & {\qipa — (—)} \\
					\textbf{A3} & \eng{dead} & {\qipa үле (—)} & {\qipa үле (—)} & {\qipa — (—)} \\
					\textbf{A4} & \eng{basic} & {\qipa төп (—)} & {\qipa — (—)} & {\qipa — (—)} \\
					\bottomrule
				\end{tabular}
			}\\
			This phenomenon is of particular interest because native-speaker intuitions on the morphological limitations of a given adjective can often be more restrictive than the range of possible uses found in a large text corpus.
	}

	\headerbox{Ongoing and future work}{name=nedostatki,column=2,below=tuvanvelars}{
		\begin{itemize}
			\item Disambiguation, more stems, clean up transducers
			\item Machine translation between these languages
			\item Bring other Kypchak transducers to comparable performance:\\
			Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
			\item Other Turkic lgs:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
		\end{itemize}
	}

	\headerbox{Further information}{name=getting,column=2,below=nedostatki}{
		\begin{itemize}
			\item Part of \textbf{Apertium Turkic} project:\\
			\url{http://wiki.apertium.org/wiki/Apertium\_Turkic}
			\item Transducers available \textbf{live} at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
			\item \textbf{Source code} available from Apertium's SVN repo
			%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}
				\item Turkic RBMT \textbf{mailing list} (>25 subscribers):\\
			\texttt{apertium-turkic@lists.sourceforge.net}\\
			\vspace{-0.5ex}Feel free to post in any language!\\\vspace{-2.5ex}
			\item And feel free to contact the authors any time!
		\end{itemize}
	}

%	\headerbox{References}{name=references,column=2,below=nedostatki}{
%		%\begin{thebibliography}{1}\itemsep=-0.01em
%		\setlength{\baselineskip}{0.4em}
%	}

@@

@@

		\headerbox{References}{name=references,column=3,below=tuvanvowelharmony}{
			\begin{spacing}{0.7}
				\printbibliography[heading=none]
			\end{spacing}
			\vspace{-0.5em}
		}

@@

				\path[line] (assumptions) -- (compile);
				\path[line] (compile) -- (analyse);
				\path[line] (analyse) -- (check);
				\path[line] (check) -- node [right=0.75ex, scale=\flowchartscale, align=left] {unanalysed\\forms} (missingstem);
				\path[line] (missingstem) -| node [below=1ex, near start, scale=\flowchartscale] {yes} (addstem);
				\path[line] (addstem) -- (compile);
				\path[line] (missingstem) -- node [below=1ex, midway, scale=\flowchartscale] {no} (phonormorph);
				\path[line] (phonormorph) -- node [below=1ex, near start, scale=\flowchartscale] {phon} (updatephon);
				\path[line] (phonormorph) |- node [below=1ex, near end, scale=\flowchartscale] {morph} (updatemorph);
				\path[line] (updatephon) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (phonbetter) ;
				\path[line] (updatemorph) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (morphbetter) ;
				\path[line] (phonbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.6,0) |- (newgeneralisation.east) ;
				\path[line] (morphbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.6,1) |- (newgeneralisation.east) ;
				\path[line] (phonbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,0.5) -| (updatephon.north) ;
				\path[line] (morphbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,1.5) -| (updatemorph.north) ;
				\path[line] (newgeneralisation) -- (compile);

>>>>

\tikzset{
  every overlay node/.style={
    draw=none,fill=none,rounded corners,anchor=north west,
  },
}
% Usage:
% \tikzoverlay at (-1cm,-5cm) {content};
% or
% \tikzoverlay[text width=5cm] at (-1cm,-5cm) {content};
\def\tikzoverlay{%
   \tikz[baseline,overlay]\node[every overlay node]
}%

@@

%\usepackage{natbib}
\usepackage[backend=biber,citestyle=apa,maxcitenames=2,style=authoryear,maxbibnames=99,uniquelist=false,uniquename=false]{biblatex}
\bibliography{poster}
\renewcommand*{\nameyeardelim}{\addcomma\space}
\renewcommand*{\bibfont}{\footnotesize}
\renewcommand*{\intitlepunct}{\space}

@@

	\headerbox{Morphological Transducers}{name=morphtrans,below=morphdesc}{
		\htwo{Morphological transducers}
		\begin{itemize}
			\item Efficient (in speed \& size) models of a language's morphology
			\item Take a surface form, and produce valid lexical form(s)
			\item Take a lexical form, and produce valid surface form(s)\vspace{-0.1ex}\\
		%\end{itemize}
		%\scalebox{0.98}{
			алдым~~~\hilitetwo{↔}~~~\texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}},~~\texttt{алд\tag{n}\tag{px1sg}\tag{nom}}
		%}
		\end{itemize}

@@

		\headerbox{Example: Tuvan velar elision}{name=tuvanvelars,column=1,span=2}{
			\begin{multicols}{2}
				\htwo{Previous descriptions}

@@

			\columnbreak{}
				\htwo{Unanalysed forms before implementation}
					\begin{itemize}
						\item \hilitetwo{инээ}, \hiliteone{өгү}, \hilitetwo{өө}, \hilitetwo{соо}, \hiliteone{чаңы}
					\end{itemize}\vspace{-0.3ex}
					\begin{minipage}[t]{0.5\textwidth}
					\outputex{
						\begin{multicols}{4}
							1590 соонда\vphantom{ү}\\
							610 те\vphantom{ү}\\
							153 өөнүң\vphantom{ү}\\
							48 соондан\vphantom{ү}\\
							46 өөм\vphantom{ү}\\
							20 соон\vphantom{ү}\\
							18 өө\vphantom{ү}\\
							17 өөнге\vphantom{ү}\\
							9 соондагы\vphantom{ү}\\
							8 өөн\vphantom{ү}\columnbreak\\
							6 инээниң\vphantom{ү}\\
							6 соонче\vphantom{ү}\\
							6 соонга\vphantom{ү}\\
							3 инээм\vphantom{ү}\\
							3 өөнден\vphantom{ү}\\
							3 өөнде\vphantom{ү}\\
							3 өөмнүң\vphantom{ү}\\
							2 инээн\vphantom{ү}\\
							2 инээ\vphantom{ү}\\
							2 өөнче\vphantom{ү}\columnbreak\\
							2 өөңерге\vphantom{ү}\\
							2 өөңер\vphantom{ү}\\
							2 өөмнү\vphantom{ү}\\
							2 өөвүстен\vphantom{ү}\\
							2 өөвүске\vphantom{ү}\\
							2 өөвүс\vphantom{ү}\\
							1 инээңни\vphantom{ү}\\
							1 инээмниң\vphantom{ү}\\
							1 Инээм\vphantom{ү}\\
							1 инээвистиң\vphantom{ү}\\
							1 өөңнүң\vphantom{ү}\\
							1 өөң\vphantom{ү}\\
							1 өөвүсче\vphantom{ү}\\
							1 өөвүстүң\vphantom{ү}\\
							1 өөвүстү\vphantom{ү}\\
							1 өөвүсте
						\end{multicols}}
					\end{minipage}

@@

	\headerbox{Ongoing and future work}{name=nedostatki,column=2,below=tuvanvelars}{
		%\begin{itemize}
			%\item Disambiguation, more stems, clean up transducers
			%\item Machine translation between these languages
			%\item Bring other Kypchak transducers to comparable performance:\\
			%Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
			%\item Other Turkic lgs:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
		%\end{itemize}
	}

	\headerbox{Further information}{name=getting,column=2,below=nedostatki}{
		\begin{itemize}
			\item Part of \textbf{Apertium Turkic} project:\\
			\url{http://wiki.apertium.org/wiki/Apertium\_Turkic}
			\item Transducers available \textbf{live} at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
			\item \textbf{Source code} available from Apertium's SVN repo
			%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}
				\item Turkic RBMT \textbf{mailing list} (>25 subscribers):\\
			\texttt{apertium-turkic@lists.sourceforge.net}\\
			\vspace{-0.5ex}Feel free to post in any language!\\\vspace{-2.5ex}
			\item And feel free to contact the authors any time!
		\end{itemize}
		\tikzoverlay[text width=0.8in] at (2.85in, 0.77in) {
			\includegraphics[height=0.8in]{qrcode}
		};
	}

@@

	\headerbox{Example: Categorisation}{name=adjectives,column=3,below=tuvanvowelharmony}{
		\htwo{Unrestricted 0-derivation}
		\begin{itemize}
			\item Common overgeneralisation of adjective morphology:
			\begin{itemize}
				\item all adjectives may act like nouns (i.e.\ take noun morphology)
				\item all adjectives may be used as is as adverbs
			\end{itemize}
			\item Commonly reported in Turkic language grammars
		\end{itemize}
		%Somfai Kara (2002, pp. 28-29):\ \includegraphics[width=0.53\textwidth]{img/kara28}\\
		\textcite[28-29]{somfaikara02}:\ \includegraphics[width=0.53\textwidth]{img/kara28}\\
		\-\hfill{}\includegraphics[width=0.8\textwidth]{img/kara29}

		\htwo{Attested system}
			\begin{itemize}
				\item Some adjectives may not be substantivised
				\item Some adjectives may not be used adverbially
				\item Some adjectives do not have comparative forms
				\item …a range of “adjective classes” in most Turkic languages
			\end{itemize}
		\htwo{Implementation:\ proper categorisation}
			\begin{itemize}
				\item only correct forms are analysed and generated
			\end{itemize}

			\scalebox{0.83}{
				\begin{tabular}{lllll}
					\toprule
					\textbf{Type} & \textbf{Gloss} & {\small \texttt{\tag{adj}(\tag{comp})}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{subst}}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{advl}}} \\
					\midrule
					\textbf{A1} & \eng{good} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} \\
					\textbf{A2} & \eng{old} & {\qipa иске (искерәк)} & {\qipa иске (искерәк)} & {\qipa — (—)} \\
					\textbf{A3} & \eng{dead} & {\qipa үле (—)} & {\qipa үле (—)} & {\qipa — (—)} \\
					\textbf{A4} & \eng{basic} & {\qipa төп (—)} & {\qipa — (—)} & {\qipa — (—)} \\
					\bottomrule
				\end{tabular}
			}\\
			\begin{itemize}
				\item Interestingly, native-speaker intuitions on the morphological limitations of a given adjective can often be more restrictive than the range of possible uses found in a large text corpus.
			\end{itemize}
	}

@@

		\headerbox{References}{name=references,column=1,below=tuvanvelars}{
			\begin{spacing}{0.75}
				\printbibliography[heading=none]
			\end{spacing}
			\vspace{-0.3em}
		}

@@

				\path[line] (assumptions) -- node [right=0.75ex, align=left, scale=\flowchartscale] {preliminary\\development} (compile);
				\path[line] (compile) -- (analyse);
				\path[line] (analyse) -- (check);
				\path[line] (check) -- node [right=0.75ex, scale=\flowchartscale, align=left] {unanalysed\\forms} (missingstem);
				\path[line] (missingstem) -| node [below=1ex, near start, scale=\flowchartscale] {yes} (addstem);
				\path[line] (addstem) -- (compile);
				\path[line] (missingstem) -- node [below=1ex, midway, scale=\flowchartscale] {no} (phonormorph);
				\path[line] (phonormorph) -- node [below=1ex, near start, scale=\flowchartscale] {phon} (updatephon);
				\path[line] (phonormorph) |- node [below=1ex, near end, scale=\flowchartscale] {morph} (updatemorph);
				\path[line] (updatephon) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (phonbetter) ;
				\path[line] (updatemorph) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (morphbetter) ;
				\path[line] (phonbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.6,0) |- (newgeneralisation.east) ;
				\path[line] (morphbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.6,1) |- (newgeneralisation.east) ;
				\path[line] (phonbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,0.5) -| (updatephon.north) ;
				\path[line] (morphbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,1.5) -| (updatemorph.north) ;
				\path[line] (newgeneralisation) -- (compile);
